---
title: 'Assignment 1: Zones and Networks'
author: "Carole Voulgaris"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These examples are for the Buffalo-Cheektowaga-Niagara Falls, NY metropolitan statistical area (FIPS code 15380)

# Step 0: Load libraries

```{r, message = FALSE}
options(java.parameters = "-Xmx2G")

library(tidyverse)
library(tidytransit)
library(tigris)
library(r5r)
library(osmdata)
library(sf)
library(ggthemes)
library(leaflet)
library(htmlwidgets)
library(gridExtra)
library(ggpubr)
```



# Step 1: Get Zones

```{r, results='hide', message=FALSE}
zones <- tracts(state = "NY", 
                county = c("Erie", "Niagara")) %>%
  st_transform("WGS84")
```

I'll create a quick leaflet map to view in my browser to see these zones.

```{r}
network_map <- leaflet(zones) %>%
addProviderTiles("Stamen.TonerLite") %>%
 addPolygons(fillColor = "blue", color = "lightblue", weight = 1,
              highlightOptions = highlightOptions(fillColor = "gray", fillOpacity = 0.1),
              label = zones$GEOID,
              popup = zones$GEOID) 

saveWidget(network_map, file="network_map.html")
```

When I view the map in my browser window, I can see that two zones (36029990000 and 36063990000) are entirely water. I'll delete these from the layer.

```{r}
zones <- zones %>%
  filter(GEOID != "36029990000",
         GEOID != "36063990000")
```

# Step 2: Get street network

```{r}
buffalo_bbox <- st_bbox(zones)

q <- opq(bbox = buffalo_bbox) %>%
  add_osm_feature(key = 'highway')

q %>% osmdata_xml(file = 'networks/buffalo_network.osm') 

```

You'll need to convert the `*.osm` file to a `*.pbf` using the [osmconvert program](https://wiki.openstreetmap.org/wiki/Osmconvert){target="_blank"}. You can find an executable you can run on Windows in the networks directory of this Repo. If you are on a Mac, ...

# Step 4: Get transit network

```{r}
download.file(feedlist[feedlist$t == "NFTA GTFS",]$url_d, "networks/NFTA_GTFS.zip")
```

# Map everything

```{r, message=FALSE, warning=FALSE}
r5r_core <- setup_r5("networks", verbose = FALSE)

streets <- street_network_to_sf(r5r_core) 
street_lines <- streets$edges[zones,]

transit <- transit_network_to_sf(r5r_core)
transit_routes <- transit$routes

```

```{r}
ggplot(street_lines) +
  geom_sf(size = 0.01, color = "gray") +
  geom_sf(data = transit_routes,
          size = 1,
          color = "red") +
  geom_sf(data = zones, fill = NA, color = "blue", alpha = 0.5) +
  theme_map()

#stop_r5(r5r_core)
```


  
Based on the maps, I can see that there  
  

# Step 5: Driving skim

Calculate a matrix of travel times

```{r, warning=FALSE, message=FALSE, results='hide'}
zone_centroids <- st_centroid(zones) %>%
  rename(id = GEOID) %>%
  select(id)

drive_skim <- travel_time_matrix(r5r_core, 
                   origins = zone_centroids,
                   destinations = zone_centroids,
                   mode = "CAR",
                   verbose = FALSE)

drive_skim_mat <- drive_skim %>%
  pivot_wider(names_from = toId, values_from = travel_time, names_prefix = "to_")
```

Select a couple of zones (one centrally located (36029016500), two near the edges (36029940000 is south, 36063024102 is north)) to illustrate travel times.

```{r}
drive_skim_zones <- zones %>%
  select(GEOID) %>%
  rename(fromId = GEOID) %>%
  left_join(drive_skim_mat) %>%
  rename(GEOID = fromId)

SW_skims <- ggplot(drive_skim_zones) +
  geom_sf(aes(fill = to_36029940000), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029940000"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 110)) +
  theme_map() +
  theme(legend.position = "none")

NE_skims <- ggplot(drive_skim_zones) +
  geom_sf(aes(fill = to_36063024102), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36063024102"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 110)) +
  theme_map() +
  theme(legend.position = "none")

central_skims <- ggplot(drive_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029016500"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 110)) +
  theme_map() +
  theme(legend.position = "none")

legend_skims <- ggplot(drive_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  scale_fill_viridis_c(limits = c(0, 110),
                       name = "Travel time from\nhighlighted zone\n(minutes)") 

legend <- as_ggplot(get_legend(legend_skims))

grid.arrange(SW_skims, NE_skims, central_skims, legend, ncol=4)
```


# Step 6: Cycling skim


```{r}
bike_skim <- travel_time_matrix(r5r_core, 
                   origins = zone_centroids,
                   destinations = zone_centroids,
                   mode = "BICYCLE",
                   max_trip_duration = 600,
                   verbose = FALSE)

bike_skim_mat <- bike_skim %>%
  pivot_wider(names_from = toId, values_from = travel_time, names_prefix = "to_")
```

Select a couple of zones (one centrally located (36029016500), two near the edges (36029940000 is south, 36063024102 is north)) to illustrate travel times.

```{r}
bike_skim_zones <- zones %>%
  select(GEOID) %>%
  rename(fromId = GEOID) %>%
  left_join(bike_skim_mat) %>%
  rename(GEOID = fromId)

SW_skims_bike <- ggplot(bike_skim_zones) +
  geom_sf(aes(fill = to_36029940000), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029940000"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

NE_skims_bike <- ggplot(bike_skim_zones) +
  geom_sf(aes(fill = to_36063024102), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36063024102"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

central_skims_bike <- ggplot(bike_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029016500"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

legend_skims <- ggplot(bike_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  scale_fill_viridis_c(limits = c(0, 600),
                       name = "Travel time from\nhighlighted zone\n(minutes)") 

legend <- as_ggplot(get_legend(legend_skims))

grid.arrange(SW_skims_bike, NE_skims_bike, central_skims_bike, legend, ncol=4)
```

# Step 7: Walking skim

```{r}
walk_skim <- travel_time_matrix(r5r_core, 
                   origins = zone_centroids,
                   destinations = zone_centroids,
                   mode = "WALK",
                   max_trip_duration = 600,
                   verbose = FALSE)

walk_skim_mat <- walk_skim %>%
  pivot_wider(names_from = toId, values_from = travel_time, names_prefix = "to_")
```

Select a couple of zones (one centrally located (36029016500), two near the edges (36029940000 is south, 36063024102 is north)) to illustrate travel times.

```{r}
walk_skim_zones <- zones %>%
  select(GEOID) %>%
  rename(fromId = GEOID) %>%
  left_join(walk_skim_mat) %>%
  rename(GEOID = fromId)

SW_skims_walk <- ggplot(walk_skim_zones) +
  geom_sf(aes(fill = to_36029940000), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029940000"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

NE_skims_walk <- ggplot(walk_skim_zones) +
  geom_sf(aes(fill = to_36063024102), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36063024102"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

central_skims_walk <- ggplot(walk_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029016500"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 600)) +
  theme_map() +
  theme(legend.position = "none")

legend_skims <- ggplot(walk_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  scale_fill_viridis_c(limits = c(0, 600),
                       name = "Travel time from\nhighlighted zone\n(minutes)") 

legend <- as_ggplot(get_legend(legend_skims))

grid.arrange(SW_skims_walk, NE_skims_walk, central_skims_walk, legend, ncol=4)
```

# Step 8: Transit skim

```{r}
transit_skim <- travel_time_matrix(r5r_core, 
                   origins = zone_centroids,
                   destinations = zone_centroids,
                   mode = "TRANSIT",
                   max_trip_duration = 600,
                   max_walk_dist = 1000,
                   verbose = FALSE)

transit_skim_mat <- transit_skim %>%
  pivot_wider(names_from = toId, values_from = travel_time, names_prefix = "to_")
```

Select a couple of zones (one centrally located (36029016500), two near the edges (36029940000 is south, 36063024102 is north)) to illustrate travel times.

```{r}
transit_skim_zones <- zones %>%
  select(GEOID) %>%
  rename(fromId = GEOID) %>%
  left_join(transit_skim_mat) %>%
  rename(GEOID = fromId)

SW_skims_transit <- ggplot(transit_skim_zones) +
  geom_sf(aes(fill = to_36029940000), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029940000"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 160)) +
  theme_map() +
  theme(legend.position = "none")

NE_skims_transit <- ggplot(transit_skim_zones) +
  geom_sf(aes(fill = to_36063024102), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36063024102"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 160)) +
  theme_map() +
  theme(legend.position = "none")

central_skims_transit <- ggplot(transit_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  geom_sf(data = zones %>% filter(GEOID=="36029016500"),
          fill = NA, color = "white", size = 1) +
  scale_fill_viridis_c(limits = c(0, 160)) +
  theme_map() +
  theme(legend.position = "none")

legend_skims <- ggplot(transit_skim_zones) +
  geom_sf(aes(fill = to_36029016500), color = NA) +
  scale_fill_viridis_c(limits = c(0, 160),
                       name = "Travel time from\nhighlighted zone\n(minutes)") 

legend <- as_ggplot(get_legend(legend_skims))

grid.arrange(SW_skims_transit, NE_skims_transit, central_skims_transit, legend, ncol=4)
```


```{r}
stop_r5()
```